Dealing with Stochastic Volatility in Time Series Using the \proglang{R} Package \pkg{stochvol}

### 2015-07-15 inst/doc/article.Rnw
version 1.2.0

%\VignetteIndexEntry{Dealing with Stochastic Volatility in Time Series Using the R Package stochvol}
%\VignetteKeyword{Bayesian inference, Markov chain Monte Carlo (MCMC), auxiliary mixture sampling, ancillarity-sufficiency interweaving strategy (ASIS), state-space model, heteroskedasticity, financial time series}

@@

%% almost as usual
\author{Gregor Kastner\\WU Vienna University of Economics and Business}
\title{Dealing with Stochastic Volatility in Time Series Using the \proglang{R} Package \pkg{stochvol}}
%\Shorttitle{The \proglang{R} package \pkg{stochvol} (DRAFT VERSION)} %% a short title (if necessary)
\Plaintitle{Dealing with Stochastic Volatility in Time Series Using the R Package stochvol} %% without formatting

@@

%% an abstract and keywords
\Abstract{
The \proglang{R} package \pkg{stochvol} provides a fully Bayesian implementation of heteroskedasticity modeling within the framework of stochastic volatility.
It utilizes Markov chain Monte Carlo (MCMC) samplers to conduct inference by obtaining draws from the posterior distribution of parameters and latent variables which can then be used for predicting future volatilities. The package can straightforwardly be employed as a stand-alone tool; moreover, it allows for easy incorporation into other MCMC samplers. The main focus of this paper is to show the functionality of \pkg{stochvol}. In addition, it provides a brief mathematical description of the model, an overview of the sampling schemes used, and several illustrative examples using exchange rate data.
}
\Keywords{Bayesian inference, Markov chain Monte Carlo (MCMC), auxiliary mixture sampling, ancillarity-sufficiency interweaving strategy (ASIS), state-space model, heteroskedasticity, financial time series}
%\Plainkeywords{Bayesian inference, Markov chain Monte Carlo (MCMC), auxiliary mixture sampling, ancillarity-sufficiency interweaving strategy (ASIS), state-space model, heteroskedasticity, financial time series}
%% at least one keyword must be supplied

@@

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Gregor Kastner\\
  Institute for Statistics and Mathematics\\
  Department of Finance, Accounting and Statistics\\
  WU Vienna University of Economics and Business\\
%  Building D4, Level 4\\
  Welthandelsplatz 1\\
  1020 Vienna, Austria\\
%  Telephone: +43/1/31336-5593\\
%  Fax: +43/1/31336-90-5593\\
  E-mail: \email{gregor.kastner@wu.ac.at}\\
  URL: \url{http://www.wu.ac.at/statmath/en/faculty_staff/faculty/gkastner/}
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/512/507-7103
%% Fax: +43/512/507-2851

@@

\section*{Preface}
This vignette corresponds to a forthcoming article of the same name in the Journal of Statistical Software. At the time of this writing, the vignette and the published article coincide; nevertheless, this version might receive minor updates over time. To cite, please use \cite{kas:dea}. Further information about citing \pkg{stochvol} can be obtained in \proglang{R} by installing the package, e.g., through \code{install.packages("stochvol")}, and typing \code{citation("stochvol")}.

@@

<<usd6, fig=TRUE, include=FALSE, width=8, height=3.5>>=
par(mfrow = c(1, 3))
paradensplot(res)
@

@@

<<homo, echo=FALSE, fig=TRUE, include=FALSE, width=7.5, height=6>>=
par(mar = c(3.1, 1.8, 1.9, .5), mgp = c(1.8, .6, 0))
plot(coda::mcmc(draws1))
@

<<eval=FALSE>>=
plot(coda::mcmc(draws1))
@

@@

<<eval=FALSE>>=
plot(coda::mcmc(draws2[, 4:7]))
@

<<hetero, fig=TRUE, include=FALSE, echo=FALSE, results=hide, width=7.5, height=8>>=
par(mar = c(3.1, 1.8, 1.9, .5), mgp = c(1.8, .6, 0))
plot(coda::mcmc(draws2selection[, 1:4]))
@

@@

\begin{table}[!ht]
 \begin{small}
 \centering
 \begin{tabular}{lccc}
  \hline
  \hline
 Currency & SV & GARCH & homoskedastic \\
  \hline
  AUD & 7938 & 7890 & 7554 \\
  CAD & 7851 & 7844 & 7714 \\
  CHF & 9411 & 9337 & 8303 \\
  CZK & 9046 & 8995 & 8659 \\
  DKK\tablefootnote{For exchange rates of the Danish krone with respect to the euro, we analyze \code{1000*log(exrates\$DKK)}, i.e., per mille (\textperthousand) log prices. This way, we avoid convergence issues when obtaining starting values for the MCMC sampler from \pkg{fGarch} that appear otherwise because the krone is pegged very closely to the euro.}
  & 1473 & 1366 & 1178 \\
  GBP & 8568 & 8549 & 8218 \\
  HKD & 7907 & 7897 & 7728 \\
  IDR & 7697 & 7662 & 7269 \\
  JPY & 7607 & 7586 & 7280 \\
  KRW & 7766 & 7749 & 7188 \\
  MXN & 7536 & 7507 & 7055 \\
  MYR & 8082 & 8064 & 7928 \\
  NOK & 8648 & 8616 & 8331 \\
  NZD & 7619 & 7587 & 7440 \\
  PHP & 7890 & 7862 & 7654 \\
  PLN & 8126 & 8094 & 7727 \\
  RON & 9011 & 8880 & 8255 \\
  RUB & 8664 & 8617 & 8146 \\
  SEK & 9110 & 9101 & 8648 \\
  SGD & 8540 & 8529 & 8308 \\
  THB & 7867 & 7844 & 7692 \\
  \hline
  USD & 7878 & 7868 & 7701 \\
  USD [AR(0)]\tablefootnote{These results refer to the analysis of log returns by means of a intercept-only model, i.e., AR(0).}
  & 7876 & 7865 & 7699 \\
  \hline
  \hline
 \end{tabular}
 \caption{Final cumulative predictive log likelihoods for AR(1) models with different error assumptions, applied to the logarithm of several EUR exchange rates.% included in \pkg{stochvol}'s dataset \code{exrates}.
 SV is strongly favored in all cases.}
 \label{exrates}
 \end{small}
\end{table}

>>>>

%\VignetteIndexEntry{Dealing with Stochastic Volatility in Time Series Using the R Package stochvol}

%\VignetteKeyword{Bayesian inference}
%\VignetteKeyword{Markov chain Monte Carlo (MCMC)}
%\VignetteKeyword{heteroskedasticity}
%\VignetteKeyword{SV}
%\VignetteKeyword{GARCH}
%\VignetteKeyword{forecasting}
%\VignetteKeyword{predictive density}
%\VignetteKeyword{predictive Bayes factor}
%\VignetteKeyword{exchange rates}

@@

%% almost as usual
\author{Gregor Kastner\\WU Vienna University of Economics and Business}
\title{Dealing with Stochastic Volatility in Time Series Using the \proglang{R} Package \pkg{stochvol}}
\Plaintitle{Dealing with Stochastic Volatility in Time Series Using the R Package stochvol} %% without formatting

@@

%% an abstract and keywords
\Abstract{
The \proglang{R} package \pkg{stochvol} provides a fully Bayesian implementation of heteroskedasticity modeling within the framework of stochastic volatility.
It utilizes Markov chain Monte Carlo (MCMC) samplers to conduct inference by obtaining draws from the posterior distribution of parameters and latent variables which can then be used for predicting future volatilities. The package can straightforwardly be employed as a stand-alone tool; moreover, it allows for easy incorporation into other MCMC samplers. The main focus of this paper is to show the functionality of \pkg{stochvol}. In addition, it provides a brief mathematical description of the model, an overview of the sampling schemes used, and several illustrative examples using exchange rate data.
}
\Keywords{Bayesian inference, Markov chain Monte Carlo (MCMC), heteroskedasticity, SV, GARCH, forecasting, predictive density, predictive Bayes factor, exchange rates}
%% at least one keyword must be supplied

@@

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Gregor Kastner\\
  Institute for Statistics and Mathematics\\
  Department of Finance, Accounting and Statistics\\
  WU Vienna University of Economics and Business\\
  Welthandelsplatz 1 / D4 / level 4\\
  1020 Vienna, Austria\\
%  Telephone: +43/1/31336-5593\\
%  Fax: +43/1/31336-90-5593\\
  E-mail: \email{gregor.kastner@wu.ac.at}\\
  URL: \url{http://www.wu.ac.at/statmath/en/faculty_staff/faculty/gkastner/}
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/512/507-7103
%% Fax: +43/512/507-2851

@@

\section*{Preface}
This vignette corresponds to a forthcoming article of the same name in the Journal of Statistical Software. At the time of this writing, the vignette and the published article coincide; nevertheless, this version might receive minor updates as time goes by. To cite, please use \cite{kas:dea}. Further information about citing \pkg{stochvol} can be obtained in \proglang{R} by installing the package, e.g., through \code{install.packages("stochvol")}, and calling \code{citation("stochvol")}.

@@

<<usd6, fig=TRUE, include=FALSE, width=8, height=3.5>>=
par(mfrow = c(1, 3))
paradensplot(res, showobs = FALSE)
@

@@

<<homo, echo=FALSE, fig=TRUE, include=FALSE, width=7.5, height=6>>=
par(mar = c(3.1, 1.8, 1.9, .5), mgp = c(1.8, .6, 0))
plot(coda::mcmc(draws1), show.obs = FALSE)
@

<<eval=FALSE>>=
plot(coda::mcmc(draws1), show.obs = FALSE)
@

@@

<<eval=FALSE>>=
plot(coda::mcmc(draws2[, 4:7]), show.obs = FALSE)
@

<<hetero, fig=TRUE, include=FALSE, echo=FALSE, results=hide, width=7.5, height=8>>=
par(mar = c(3.1, 1.8, 1.9, .5), mgp = c(1.8, .6, 0))
plot(coda::mcmc(draws2selection[, 1:4]), show.obs = FALSE)
@

@@

\begin{table}[!ht]
 \begin{small}
 \centering
 \begin{tabular}{lccc}
  \hline
  \hline
 Currency & SV & GARCH & homoskedastic \\
  \hline
  AUD & 7938 & 7890 & 7554 \\
  CAD & 7851 & 7844 & 7714 \\
  CHF & 9411 & 9337 & 8303 \\
  CZK & 9046 & 8995 & 8659 \\
  DKK\tablefootnote{For exchange rates of the Danish krone with respect to the euro, we analyze \code{1000*log(exrates\$DKK)}, i.e., per mille (\textperthousand) log prices. This way, we avoid convergence issues when obtaining starting values for the MCMC sampler from \pkg{fGarch} that appear otherwise because the krone is pegged very closely to the euro.}
  & 1473 & 1366 & 1178 \\
  GBP & 8568 & 8549 & 8218 \\
  HKD & 7907 & 7897 & 7728 \\
  IDR & 7697 & 7662 & 7269 \\
  JPY & 7607 & 7586 & 7280 \\
  KRW & 7766 & 7749 & 7188 \\
  MXN & 7536 & 7507 & 7055 \\
  MYR & 8082 & 8064 & 7928 \\
  NOK & 8648 & 8616 & 8331 \\
  NZD & 7619 & 7587 & 7440 \\
  PHP & 7890 & 7862 & 7654 \\
  PLN & 8126 & 8094 & 7727 \\
  RON & 9011 & 8880 & 8255 \\
  RUB & 8664 & 8617 & 8146 \\
  SEK & 9110 & 9101 & 8648 \\
  SGD & 8540 & 8529 & 8308 \\
  THB & 7867 & 7844 & 7692 \\
  \hline
  USD & 7878 & 7868 & 7701 \\
  USD [AR(0)]\tablefootnote{These results refer to the analysis of log returns by means of a intercept-only model, i.e., AR(0).}
  & 7876 & 7865 & 7699 \\
  \hline
  \hline
 \end{tabular}
 \caption{Final cumulative predictive log likelihoods for AR(1) models with different error assumptions, applied to the logarithm of several EUR exchange rates. % included in \pkg{stochvol}'s dataset \code{exrates}.
 SV is strongly favored in all cases.}
 \label{exrates}
 \end{small}
\end{table}
